Consensus in the Browser using WebRTC

### 2020-06-12 paper/main.tex
more paper progress

\documentclass[11pt,twocolumn]{article}
\usepackage{graphicx}
\usepackage{biblatex}
\addbibresource{main.bib}

@@

We believe there is significant potential in using less centralized models to build programs that meet expectations while respecting the customer. However, the decentralized approach poses new challenges around performance and predictability that are not typically as challenging in more controlled environments.

\section{Choosing Raft}
Achieving consensus is a fundamental problem in the space of Distributed Computing. As a result, many distributed consensus algorithms exist, including the pervasive Paxos, Viewstamped Replication, and the newer Proof of Work approach. When considering algorithms for this project, we sought out these characteristics:
\begin{itemize}
    \item \textbf{Understandability}: This project was conceived primarily as a learning exercise for the author, who has minimal prior distributed systems experience. Therefore, the feasibility of implementing the consensus algorithm in the time allotted was a driving concern.
    \item \textbf{Low Latency}: A primary objective of this project was to use the consensus algorithm in an interactive application. To provide a responsive experience, low latency was important to consider.
\end{itemize}

@@

When a peer seeks to establish an WebRTC connection, it first gathers information on how other peers may route packets to it. This information, known as an ICE (Interactive Connectivity Establishment) candidate, is then sent to the target peer via the known relay server. Upon receiving an ICE candidate, the receiving registers that information, seeks out its own ICE candidates, and relays them back to the sending peer. At this point, both peers have the necessary knowledge to establish a direct peer-to-peer connection.

@@

Once connectivity and session information are exchanged, the two peers have all the necessary information to begin sending information directly to each other.

@@

\subsection{Building a Cluster}
In order for our application to work without each peer having predefined knowledge of each node in the cluster, we designed a simple protocol that allowed peers to discover and connect to each other peer in the cluster as they come online. The protocol defines four messages: "register", "discover", "offer" and "answer", as described in Figure \ref{fig:cluster}.

@@

When a node comes online and seeks to become part of the cluster, it issues a "register" call to our signaling server, which records the node's uuid as well as a socket for future communication with that node.

Every time a new node becomes registered, our signaling server broadcasts a "discover" message to every registered node prior to that point with the uuid of the newly registered node.

Upon receiving a "discover" message from the signaling server, the node then creates an SDP and forwards that via the signaling server to each newly discovered node.

Finally, upon receiving an "offer" message from the signaling server, the receiving node forwards "answer" message along with its own SDP via the signaling server to the originator of the "offer" message. At this point, every node in the cluster has the necessary information to establish peer to peer connections with each other and we can begin our Raft session.

@@

\subsection{Raft}
To understand the intricacies of the Raft consensus algorithm, we built a from-scratch implementation in Typescript. For the sake of testing, we began by implementing Raft in a single process running in Node. This allowed us to easily construct scenarios described in the original Raft paper and quickly identify mistakes in our implementation. Once we felt that the fundamentals of our implementation were correct, we extracted our in-process RPC implementation and replaced it with RPC conducted over WebRTC. This approach has the added benefit of making our implementation somewhat platform agnostic. With minimal effort, we could provide RPC implementations over arbitrary transports and have versions of Raft that work in a browser using HTTP, or in a headless context via Node.

@@

@@

\subsection{Scalability}
To evaluate the latency of our implementation, we utilized Google's Puppeteer browser automation tooling to simulate Raft followers sending append RPCs sequentially for a span of 10 seconds. Measuring the number of sequential messages committed while varying the cluster size provided insight into how the implementation scales with larger clusters. We conducted 10 trials for each cluster size.\\

\begin{figure}[t]
\centering
\includegraphics[scale=0.4]{performance}
\caption{Latency of system over cluster size. Conducted in Google Chrome on a 2016 MacBook Pro}
\end{figure}

@@

\subsection{Predictability}
Many consensus-driven applications operate in highly controlled environments with certain guarantees on hardware and network performance. Achieving consensus on end-customer devices means that we cannot operate with any assumptions about the customer's computing conditions. In this context, the Raft consensus algorithm suffers, as it relies on tuned timeouts to determine when to begin leader election and cluster membership changes.

Honey Badger is a consensus algorithm designed with unpredictable network conditions in mind. However, its relatively high latency when compared with Raft and PBFT makes it less ideal for low-latency applications such as chat.

@@

For these reasons, future additional work may want to evaluate the feasibility of byzantine fault tolerant algorithms in the browser context. Practical Byzantine Fault Tolerance (PBFT) from Liskov et al. demonstrates byzantine fault tolerance with acceptable latency numbers for an application requiring responsiveness, such as chat. However, PBFT still relies on a concept of weak synchrony in which the network latency does not increase exponentially. This weak synchrony may not necessarily hold in less predictable environments as described in section 6.2.

\section{Conclusion}
TODO

\printbibliography

>>>>

\documentclass[11pt,twocolumn]{article}
\usepackage{graphicx}
\usepackage{biblatex}

\addbibresource{main.bib}

@@

We believe there is significant potential in using less centralized models to build programs that meet expectations while respecting the customer. However, the decentralized approach also poses new challenges around performance and predictability that are not as difficult to overcome in controlled environments.

\section{Choosing Raft}
Achieving consensus is a fundamental problem in the space of distributed computing. As a result, many distributed consensus algorithms exist, including the pervasive Paxos \cite{paxospaper}, Viewstamped Replication \cite{vsrpaper}, and Proof of Work \cite{powpaper}. When considering algorithms for this project, we sought out these characteristics:
\begin{itemize}
    \item \textbf{Understandability}: This project was conceived primarily as a learning exercise for the author, who has minimal prior distributed systems experience. Therefore, the feasibility of implementing the consensus algorithm in the time allotted was a driving concern.
    \item \textbf{Low Latency}: A primary objective of this project was to use the consensus algorithm in an interactive application. To provide a responsive experience, low latency was important to consider.
\end{itemize}

@@

When a peer seeks to establish an WebRTC connection, it first gathers information on how other peers may route packets to it. This information, known as an ICE (Interactive Connectivity Establishment) candidate, is then sent to the target peer via the known relay server. Upon receiving an ICE candidate, the receiving peer registers that information, seeks out its own ICE candidates, and relays them back to the sending peer. At this point, both peers have the necessary knowledge to establish a direct peer-to-peer connection.

@@

Once connectivity and session information are exchanged, the two peers have all the necessary information to begin sending packets directly to each other.

@@

\subsection{Building a Cluster}
In order for our application to work without each peer having predefined knowledge of each other peer in the cluster, we designed a simple protocol that allowed peers to discover and connect to each other peer in the cluster as they come online. The protocol defines four messages: \textit{register}, \textit{discover}, \textit{offer} and \textit{answer}, as described in Figure \ref{fig:cluster}.

@@

When a node comes online and seeks to become part of the cluster, it issues a \textit{register} call to our signaling server, which records the node's uuid as well as a socket for future communication with that node.

Every time a new node becomes registered, our signaling server broadcasts a \textit{discover} message to every registered node prior to that point with the uuid of the newly registered node.

Upon receiving a \textit{discover} message from the signaling server, the node then creates an SDP and forwards that via the signaling server to each newly discovered node.

Finally, upon receiving an \textit{offer} message from the signaling server, the receiving node forwards an \textit{answer} message along with its own SDP via the signaling server to the originator of the \textit{offer} message. At this point, every node in the cluster has the necessary information to establish peer to peer connections with each other and we can begin our Raft session.

@@

\subsection{Raft}
To understand the intricacies of the Raft consensus algorithm, we built a from-scratch implementation in Typescript. For the sake of testing, we began by implementing Raft in a single process running in Node. This allowed us to easily construct scenarios described in the original Raft paper and quickly identify mistakes in our implementation. Once we felt that the fundamentals of our implementation were correct, we extracted our in-process RPC implementation and replaced it with RPC conducted over WebRTC. This approach has the added benefit of making our implementation somewhat platform agnostic. With minimal effort, we could provide RPC implementations over arbitrary transports and have versions of Raft that work in a browser using XMLHttpRequests, or in a headless context via Node and websockets.

@@

Because all communication is done directly between peers via WebRTC, no data is ever exposed to a centralized entity. Therefore, the peer-to-peer approach inherently protects a user's data better than in the centralized approach, where a user's data may become compromised as part of an attack on the managing entity.

@@

\subsection{Scalability}
To evaluate the latency of our implementation, we utilized Google's Puppeteer \cite{puppeteerwebsite} browser automation tooling to simulate Raft followers sending append RPCs sequentially for a span of 10 seconds. Measuring the number of sequential messages committed while varying the cluster size provided insight into how the implementation scales with larger clusters. We conducted 10 trials for each cluster size. The results of this evaluation can be seen in Figure \ref{fig:scale}.

\begin{figure}[t]
\centering
\includegraphics[scale=0.4]{performance}
\caption{Latency of system over cluster size. Conducted in Google Chrome on a 2016 MacBook Pro}
\label{fig:scale}
\end{figure}

@@

For these reasons, future additional work may want to evaluate the feasibility of byzantine fault tolerant algorithms in the browser context. Practical Byzantine Fault Tolerance (PBFT) \cite{pbftpaper} demonstrates byzantine fault tolerance with acceptable latency numbers for an application requiring responsiveness, such as chat. However, PBFT still relies on a concept of weak synchrony in which the network latency does not increase exponentially. This weak synchrony may not necessarily hold in less predictable environments as described in section 6.3.

\subsection{Predictability}
Many consensus-driven applications operate in highly controlled environments with certain guarantees on hardware and network performance. Achieving consensus on end-customer devices means that we cannot operate with any assumptions about the customer's computing conditions. In this context, the Raft consensus algorithm suffers, as it relies on tuned timeouts to determine when to begin leader election and cluster membership changes. For this project, the timeouts were manually chosen to reflect the author's network conditions. Future work may consider using dynamic timeouts as cluster members' network conditions change.

Honey Badger \cite{honeybadgerpaper} is a consensus algorithm designed with unpredictable network conditions in mind. However, its relatively high latency when compared with Raft and PBFT makes it less ideal for low-latency applications such as chat.

@@

\section{Conclusion}
The centralized model for building consensus-based applications is often the go-to choice, as it provides more predictability and control for the implementor. However, this approach also requires trust from the end customer and incurs management costs for the providing entities. As network availability increases and consumer devices become more powerful, we believe there is significant potential in the decentralized model for building popular, reliable, trustworthy applications.

In this paper, we analyzed the viability of this decentralized approach by using accepted technologies to build a product with high expectations on reliability and latency. We provided a from-scratch implementation of the Raft consensus algorithm that uses WebRTC to facilitate direct communication between cluster peers. We used our Raft implementation to build and evaluate a chat application with consensus-based chat history. We found that Raft and WebRTC could be used to build a product that met customer expectations without exposing customer data to a centralized entity. We also found that the decentralized environment introduced new challenges around scalability, predictability of network conditions, and trustworthy persistence of customer data.

\printbibliography[type=book]
